---
title: "Certified Robustness Values with Sandwich Standard Errors"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Certified Robustness Values with Sandwich Standard Errors}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(RobustSandwichRV)
```

## The object being computed

For a scalar omitted confounder, let

\[
x=R^2_{D\sim Z\mid X},\qquad
y=R^2_{Y\sim Z\mid D,X},\qquad
\omega_R(Z)=\frac{se_R(\widehat\tau_f)}{se_C(\widehat\tau_f)}.
\]

The bounded calibrated value uses an observed-control magnitude $b$ and the
zero-strength synthetic-index origin $\omega_{R,00}$:

\[
\mathcal W_R(b)=
[\omega_{R,00}e^{-b},\omega_{R,00}e^b].
\]

It finds the smallest $\max(x,y)$ among attainable confounders that change the
robust test decision and have $\omega_R(Z)\in\mathcal W_R(b)$.

The software first computes an analytical protected lower bound at the adverse
band endpoint. It then either certifies equality constructively or returns the
best independently refitted witness as a certified upper bound.

The result records `theorem_exact` separately from `numerically_closed`. The
former requires the adverse-endpoint equality certificate; the latter only
means that a verified lower--upper interval is no wider than the stated
numerical tolerance.

## Reproducible HC1 example

```{r data}
set.seed(20260809)
n <- 120
x1 <- rnorm(n)
x2 <- rnorm(n)
d <- 0.4 * x1 - 0.2 * x2 + rnorm(n)
e <- (0.7 + 0.3 * abs(d)) * rnorm(n)
y <- 0.55 * d + 0.25 * x1 - 0.15 * x2 + e
fit <- lm(y ~ d + x1 + x2)
```

We use `x1` as an observed scalar benchmark.

```{r benchmark}
benchmark <- benchmark_from_model(fit, "d", "x1", vcov = "HC1")
benchmark
```

```{r bounded}
result <- robust_rv(
  fit,
  treatment = "d",
  vcov = "HC1",
  benchmark = benchmark,
  n_starts = 8,
  max_iterations = 20
)
result
```

`RV_lower` is analytical. A finite `RV_upper` always has an explicit
`best_witness_z` and `direct_refit_valid = TRUE`. `LOWER_BOUND_ONLY` means that
the numerical search did not find an upper witness; it is not a global
infeasibility statement.

## CR1 uses the same algorithm

```{r cr1}
cluster <- rep(seq_len(12), each = 10)
cluster_benchmark <- benchmark_from_model(
  fit, "d", "x1", vcov = "CR1", cluster = cluster
)
cluster_result <- robust_rv(
  fit,
  treatment = "d",
  vcov = "CR1",
  cluster = cluster,
  benchmark = cluster_benchmark,
  n_starts = 8,
  max_iterations = 20
)
cluster_result
```

Only the score aggregation changes: HC sums squared observation scores, while
CR first sums scores within clusters. Protected formulas, bracketing, spherical
root finding, fallback logic, and direct validation are shared.

## Point conditioning

```{r point}
point_result <- conditional_omega_rv(
  cache_from_lm(fit, "d"),
  omega_target = 1,
  vcov = "HC1",
  method = "protected_first",
  n_starts = 8,
  max_iterations = 20
)
point_result
```

Point conditioning is useful for theory and diagnostics, but it imposes one
exact signed sandwich ratio. Bounded calibration is the recommended
observed-control workflow.

## Refit the witness explicitly

```{r refit}
if (!is.null(result$best_witness_z)) {
  witness <- evaluate_confounder_reference(
    cache_from_lm(fit, "d"),
    result$best_witness_z,
    vcov = "HC1"
  )
  c(
    x = witness$r2_d,
    y = witness$r2_y,
    omega = witness$omega,
    robust_t = witness$t_robust
  )
}
```

## Scope

Version 0.1.0 supports unweighted linear regression and one-way clustering.
HC2, HC3, CR2, HAC, weighted least squares, generalized linear models, and
critical values that change with the omitted confounder are outside the current
certificate theory.
